---
title: "Tidy Tuesday Exploration 2020-10-06"
author: "Sara Stoudt"
date: "10/7/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```

Original data (with codebook) [here](Data here: https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-10-06/readme.md).

```{r}
library(dplyr)
library(ggplot2)
library(purrr)
library(kableExtra)
setwd("~/Desktop/tidytuesday/data/2020/2020-10-06")

data <- read.csv("tournament.csv", stringsAsFactors = F)
```

### Who are the best teams in 2018?

```{r}
topTeams <- data %>%
  filter(year == 2018) %>%
  arrange(desc(reg_w)) %>%
  head(5)

kable(topTeams)
```

### Have they gotten better/worse/stayed the same over time?

```{r}
ggplot(subset(data, school %in% topTeams$school), aes(year, reg_percent, col = school, group = school)) + geom_point() + geom_line() + theme_minimal(base_size = 20)
```

### What about the best teams in 1990?

```{r}
topTeamsOld <- data %>%
  filter(year == 1990) %>%
  arrange(desc(reg_w)) %>%
  head(5)

kable(topTeams)
```

```{r}
ggplot(subset(data, school %in% topTeamsOld$school), aes(year, reg_percent, col = school, group = school)) + geom_point() + geom_line() + theme_minimal(base_size = 20)
```

## Let's see one plot per school instead (these are all jumbled up).

```{r}
ggplot(subset(data, school %in% topTeams$school), aes(year, reg_percent)) + geom_point() + geom_line() + facet_wrap(~school) + theme_minimal(base_size = 25)

ggplot(subset(data, school %in% topTeamsOld$school), aes(year, reg_percent)) + geom_point() + geom_line() + facet_wrap(~school) + theme_minimal(base_size = 25)
```

### What about seed instead of winning percentage?

```{r}
ggplot(subset(data, school %in% topTeamsOld$school), aes(year, seed)) + geom_point() + geom_line() + facet_wrap(~school) + theme_minimal(base_size = 25)

ggplot(subset(data, school %in% topTeams$school), aes(year, seed)) + geom_point() + geom_line() + facet_wrap(~school) + theme_minimal(base_size = 25)
```

### How does winning championship relate to seed?

```{r}
champ <- data %>% filter(tourney_finish == "Champ")

ggplot(champ, aes(year, seed)) + geom_point() + geom_line() + theme_minimal(base_size = 25)
```

### Who are those two teams that won even though they were a 3 seed?

```{r}
champ[which(champ$seed == 3), ]

oddballs <- c("North Carolina", "Tennessee") ## 1994, 1997
```

### What does their history look like?

```{r}
ggplot(subset(data, school %in% oddballs), aes(year, reg_percent)) + geom_point() + geom_line() + geom_vline(xintercept = 1994) + geom_vline(xintercept = 1997) + facet_wrap(~school)
```

### Is there a relationship between seed and outcome?

```{r}
ggplot(subset(data, school %in% oddballs), aes(year, seed, col = tourney_finish)) + geom_point(size = 3) + facet_wrap(~school) + theme_minimal(base_size = 20)
```

### In table form how well have they done historically?

```{r}
table(subset(data, school == "North Carolina")$tourney_finish)

table(subset(data, school == "Tennessee")$tourney_finish)
```


### What about the relationship between seed and how they got into the tournament in the first place?

```{r}
ggplot(data, aes(year, seed, col = how_qual)) + geom_point() + theme_minimal(base_size = 20)
```

### Not super enlightening, but we don't actually care about year, so let' try this.

```{r}
ggplot(data, aes(seed)) + geom_histogram(alpha = .5) + facet_wrap(~how_qual) + theme_minimal(base_size = 20) ## interesting

```