---
title: "Tidy Tuesday Week 4"
author: "Sara Stoudt"
date: "4/24/2018"
output: github_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 4 - Gender differences in Australian Average Taxable Income
[RAW DATA](https://github.com/rfordatascience/tidytuesday/blob/master/data/week4_australian_salary.csv)  
[Article](http://www.womensagenda.com.au/latest/eds-blog/australia-s-50-highest-paying-jobs-are-paying-men-significantly-more/)  
[DataSource: data.gov.au](https://data.gov.au/dataset/taxation-statistics-2013-14/resource/c506c052-be2f-4fba-8a65-90f9e60f7775?inner_span=True)

#### Disparities in STEM 

**Take-aways**

1. About equal number of indivuals in scientist jobs. 
2. Many more males in engineering jobs.

(to be fair, should look into proportion of work force)

3. Rough OLS interpretation: For every dollar a woman makes in science, a man makes $1.52.
4. Rough OLS interpretation: For every dollar a woman makes in engineering, a man makes $1.26.

```{r,message=FALSE, warning=FALSE}
setwd("~/Desktop/tidytuesday/data")
aus<-read.csv("week4_australian_salary.csv")

require(ggplot2)
require(plotly) ## use to hover and see the job names
```

### Look for STEM jobs.

```{r,message=FALSE, warning=FALSE}
aus[grep("stat",aus$occupation),] ## looking for statistics
aus[grep("math",aus$occupation),] ## nope

scientist=aus[grep("scien",aus$occupation),] ## bingo
engineer=aus[grep("engineer",aus$occupation),]
```

Get things organized. Not particularly tidy, but bear with me. 

```{r}
scientistG=split(scientist,scientist$gender)
engineerG=split(engineer,engineer$gender)

names(scientistG[[1]])=paste("F",names(scientistG[[1]]),sep="")
names(scientistG[[2]])=paste("M",names(scientistG[[2]]),sep="")

names(engineerG[[1]])=paste("F",names(engineerG[[1]]),sep="")
names(engineerG[[2]])=paste("M",names(engineerG[[2]]),sep="")

scientistFull=cbind(scientistG[[1]],scientistG[[2]])
engineerFull=cbind(engineerG[[1]],engineerG[[2]])
```

### Look at number of individuals in each job

The line is y=x. If there was gender parity, we would see points lying around this line. You can hover to see the job titles.

```{r}
p <- ggplot(scientistFull, aes(x = Findividuals, y = Mindividuals, text =Moccupation)) +
  geom_point() +geom_abline(intercept = 0, slope = 1)+xlab("number of individuals")+
  ylab("average taxable income for males ($)")+ggtitle("Science Jobs")
p ## for static version on github
p <- ggplotly(p)
p
```

```{r}
p <- ggplot(engineerFull, aes(x = Findividuals, y = Mindividuals, text =Moccupation)) +
  geom_point() +geom_abline(intercept = 0, slope = 1)+xlab("number of individuals")+
  ylab("average taxable income for males ($)")+ggtitle("Engineer Jobs")
p ## for static version on github
p <- ggplotly(p)
p
```

### Look at salary

Again the line is y=x. If there was gender parity, we would see points lying around this line. You can hover to see the job titles.

```{r}
p <- ggplot(scientistFull, aes(x = Faverage_taxable_income, y = Maverage_taxable_income, text =Moccupation)) +
  geom_point() +geom_abline(intercept = 0, slope = 1)+xlab("average taxable income for females ($)")+
  ylab("average taxable income for males ($)")+ggtitle("Science Jobs")
p ## for static version on github
#p <- ggplotly(p) ## to look at job titles
#p
```

```{r}
p <- ggplot(engineerFull, aes(x = Faverage_taxable_income, y = Maverage_taxable_income, text =Moccupation)) +
  geom_point() +geom_abline(intercept = 0, slope = 1)+xlab("average taxable income for females ($)")+
  ylab("average taxable income for males ($)")+ggtitle("Engineer Jobs")
p ## for static version on github
#p <- ggplotly(p) ## to look at job titles
#p
```

### Rough Modeling

```{r}
lm(scientistG[[2]]$Maverage_taxable_income~scientistG[[1]]$Faverage_taxable_income)
```

```{r}
lm(engineerG[[2]]$Maverage_taxable_income~engineerG[[1]]$Faverage_taxable_income)
```