---
title: "Tidy Tuesday Week 15"
author: "Sara Stoudt"
date: "7/16/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 15

[Craft Beer USA](https://github.com/rfordatascience/tidytuesday/tree/master/data)

[data.world](https://data.world/)

[thrillist.com](https://www.thrillist.com/news/nation/most-craft-breweries-by-state-united-states)


```{r,warning=F,message=F}
require(readxl)
require(dplyr)
require(ggplot2)
require(stringr)
```

```{r}
setwd("~/Desktop/tidytuesday/data")
beers=read_excel("week15_beers.xlsx",sheet=1)
brewer=read_excel("week15_beers.xlsx",sheet=2)
beer=inner_join(beers,brewer,by =c("brewery_id"="id"))
```


```{r}
byState=beer %>% group_by(state) %>% summarise(numBrewer=length(unique(brewery_id)),count=n(),mabv=mean(abv,na.rm=T))
counties= map_data("county")
state=map_data("state")

stateInfo=cbind.data.frame(abb=state.abb,name=tolower(state.name))

state=inner_join(state,stateInfo,by=c("region"="name"))

all_state=inner_join(state,byState,by=c("abb"="state"))

```


This palette isn't very visually appealing, but in the spirit of beer, I'll use it anyway.

```{r}
#https://www.reddit.com/r/beer/comments/4gd24e/the_hex_colour_palette_of_beer/
beerPal=c("#F3F993","#F5F75C","#F6F513","#EAE615","#E0D01B","#D5BC26","#CDAA37","#C1963C","#BE8C3A","#BE823A","#C17A37","#BF7138","#BC6733","#B26033","#A85839","#985336","#8D4C32","#7C452D","#6B3A1E","#5D341A","#4E2A0C","#4A2727","#361F1B","#261716","#231716","#19100F","#16100F","#120D0C","#100B0A","#050B0A")  
```

## Where to Bar Crawl?

```{r}
ggplot(data=state,mapping=aes(x=long,y=lat,group=group))+geom_polygon(data=all_state, aes(fill=mabv),color="grey")+labs(fill="mabv")+scale_fill_gradientn(colors=beerPal)+theme_void()+geom_path(data=state, aes(x=long,y=lat,group=group),color="black")+ggtitle("Where is the beer strongest?")
```

A stark (and believable) difference between Nevada and Utah.

```{r}
ggplot(data=state,mapping=aes(x=long,y=lat,group=group))+
geom_polygon(data=all_state, aes(fill=numBrewer),color="grey")+labs(fill="numBrewer")+scale_fill_gradientn(colors=beerPal)+theme_void()+geom_path(data=state, aes(x=long,y=lat,group=group),color="black") +ggtitle("Where are the most breweries?")
```

Colorado maintains it's reputation.

```{r}
ggplot(data=state,mapping=aes(x=long,y=lat,group=group))+geom_polygon(data=all_state, aes(fill=count/numBrewer),color="grey")+labs(fill="beerPerBrewer")+scale_fill_gradientn(colors=beerPal)+theme_void()+
geom_path(data=state, aes(x=long,y=lat,group=group),color="black")+ggtitle("Where is the largest variety?")
```

Surprisingly Kansas is where it is at!

```{r}
beer %>% group_by(city,state) %>% summarise(count=n(),numBrewer=length(unique(brewery_id))) %>% arrange(desc(count))
beer %>% group_by(city,state) %>% summarise(count=n(),numBrewer=length(unique(brewery_id))) %>% arrange(desc(numBrewer))
```

Somebody please tell me about the hidden gem of Grand Rapids. Apparently, it is [Beer City, USA](https://www.experiencegr.com/things-to-do/beer-city/).
 

## Styles

There are too many styles, so I pick some major ones and investigate them.

```{r}
stout=beer[str_detect(beer$style,"Stout"),] 
american=beer[str_detect(beer$style,"American"),]
ipa=beer[str_detect(beer$style,"IPA"),]
```


**True American?**

```{r}
byStateA=american %>% group_by(state) %>% summarise(numBrewer=length(unique(brewery_id)),count=n(),mabv=mean(abv,na.rm=T))

all_state=inner_join(state,byStateA,by=c("abb"="state"))
  
ggplot(data=state,mapping=aes(x=long,y=lat,group=group))+geom_polygon(data=all_state, aes(fill=numBrewer),color="grey")+labs(fill="numBrewer")+scale_fill_gradientn(colors=beerPal)+theme_void()+geom_path(data=state, aes(x=long,y=lat,group=group),color="black")
  
ggplot(data=state,mapping=aes(x=long,y=lat,group=group))+geom_polygon(data=all_state, aes(fill=count/numBrewer),color="grey")+labs(fill="beerPerBrewer")+scale_fill_gradientn(colors=beerPal)+theme_void()+geom_path(data=state, aes(x=long,y=lat,group=group),color="black") 
```

Mississippi: American Beer's Hearland


**Representing my namesake**

```{r}

byStateS=stout %>% group_by(state) %>% summarise(numBrewer=length(unique(brewery_id)),count=n(),mabv=mean(abv,na.rm=T))
  
all_state=inner_join(state,byStateS,by=c("abb"="state"))
  
ggplot(data=state,mapping=aes(x=long,y=lat,group=group))+geom_polygon(data=all_state, aes(fill=numBrewer),color="grey")+labs(fill="numBrewer")+scale_fill_gradientn(colors=beerPal)+theme_void()+geom_path(data=state, aes(x=long,y=lat,group=group),color="black") 
  
  
ggplot(data=state,mapping=aes(x=long,y=lat,group=group))+geom_polygon(data=all_state, aes(fill=count/numBrewer),color="grey")+labs(fill="beerPerBrewer")+scale_fill_gradientn(colors=beerPal)+theme_void()+geom_path(data=state, aes(x=long,y=lat,group=group),color="black") 
```

What's up with some states having no stouts?!



**The controversial IPA**

```{r}
byStateI=ipa %>% group_by(state) %>% summarise(numBrewer=length(unique(brewery_id)),count=n(),mabv=mean(abv,na.rm=T))
  
all_state=inner_join(state,byStateI,by=c("abb"="state"))
  
ggplot(data=state,mapping=aes(x=long,y=lat,group=group))+geom_polygon(data=all_state, aes(fill=numBrewer),color="grey")+labs(fill="numBrewer")+scale_fill_gradientn(colors=beerPal)+theme_void()+geom_path(data=state, aes(x=long,y=lat,group=group),color="black") 
  
  
ggplot(data=state,mapping=aes(x=long,y=lat,group=group))+
geom_polygon(data=all_state, aes(fill=count/numBrewer),color="grey")+labs(fill="beerPerBrewer")+scale_fill_gradientn(colors=beerPal)+theme_void()+
geom_path(data=state, aes(x=long,y=lat,group=group),color="black") 
  
## what's up with Utah?
ut=beer[which(beer$state == "UT"),]
ut[str_detect(ut$style, "IPA"),] ## double counting
  
  
```

West Virginia and Arkansas are not into IPAs.

### Variation in ABV

Which styles have the most variation in alcohol content (of the top 20 most prevalent styles) given their average value?

```{r}
beer %>% group_by(style)%>% summarise(count=n(),coeffVarabv=mean(abv,na.rm=T)/sd(abv,na.rm=T)) %>% arrange(desc(count))  %>% head(20) %>% arrange(desc(coeffVarabv))
```


Fancy string matching for another time: match the beer style to the colors listed [here](https://www.twobeerdudes.com/beer/srm).