---
title: "Tidy Tuesday Week 3"
author: "Sara Stoudt"
date: "5/2/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(readxl)
require(dplyr)
require(ggplot2)
require(gridExtra)
require(tidyr)
require(RColorBrewer)
```

# Setup 
```{r,eval=F}
require(readxl)
require(dplyr)
require(ggplot2)
require(gridExtra)
require(tidyr)
require(RColorBrewer)
```


# Read and Clean Data 
```{r}
setwd("~/Desktop/tidytuesday/data")
gm=read_excel("global_mortality.xlsx")

gm.gathered=gather(gm,cause,percent,-country,-country_code,-year) ## want a single column for cause of death
gm.gathered$cause=as.vector(gsub(" \\(\\%\\)","",gm.gathered$cause)) ## remove (%) in causes of death
```

# Get Colors Ready 

I will want the color per cause to be the same across plots. 

The colors I use are still not perfectly distinguishable. Any suggestions?

```{r}
colorOrder=colorRampPalette( c("red","orange","yellow","green","blue","purple"))(length(unique(gm.gathered$cause)))
colorOrderShuffle=colorOrder[sample(1:length(colorOrder),length(colorOrder))]
## don't want causes close in alphabetical order to be near the same color mainly because of prevalence of
## cancer and cardiovascular diseases


colorMap=cbind.data.frame(colorOrderShuffle,
                          unique(gm.gathered$cause))
names(colorMap)=c("color","cause")
```

# Plot Function 

Here is a function to, given a country, plot the causes of death that have the top N highest average percentage and coefficients of variation across the time span. 

```{r}
makePlotTopN=function(data, Country,N){
  
  dataToUse=subset(data,country==Country) ## get country of interest
  
  
  byCause=group_by(dataToUse,cause)%>%summarise(avgPercent=mean(percent),sdPercent=sd(percent),sdPercentNorm=sd(percent)/avgPercent) 
  ## find average and standard deviation of percentages of causes of death across the time frame
  
  byCauseM= byCause%>%arrange(desc(avgPercent))
  byCauseSD=byCause%>%arrange(desc(sdPercentNorm))
  
  toPlotM=subset(dataToUse,cause%in%byCauseM$cause[1:N]) ## get top N average
  toPlotSD=subset(dataToUse,cause%in%byCauseSD$cause[1:N]) ## get top N variability
  
  
  ## want colors to be the same across plots
  ## is there an easier way?
  toPlotM$cause=as.factor(toPlotM$cause)
  toPlotSD$cause=as.factor(toPlotSD$cause)
  
  toMergeM=as.data.frame(toPlotM$cause)
  toMergeSD=as.data.frame(toPlotSD$cause)
  names(toMergeM)=names(toMergeSD)="cause"
  
  col1=unique(merge(toMergeM,colorMap,by.x="cause",by.y="cause"))
  col2=unique(merge(toMergeSD,colorMap,by.x="cause",by.y="cause"))
  
  ## plots
  g1=  ggplot(toPlotM,aes(x=year,y=percent,color=cause))+geom_line(size=2)+
    ggtitle(paste(Country,": Top ",N," Highest Average Cause of Death"))+
    ylab("Percentage of Deaths")+xlab("Year")+scale_colour_manual(values=as.character(col1$color))
  
  g2=ggplot(toPlotSD,aes(x=year,y=percent,color=cause))+geom_line(size=2)+
    ggtitle(paste(Country,": Top ",N," Highest Coeff of Var Cause of Death"))+
    ylab("Percentage of Deaths")+xlab("Year")+scale_colour_manual(values=as.character(col2$color))
  
  
  grid.arrange(g1,g2,ncol=2)
  

  
}
```


Following the article to choose sample countries.

```{r,fig.width=14}
makePlotTopN(gm.gathered,"United States",10)
```

```{r,fig.width=14}
makePlotTopN(gm.gathered,"Germany",10)
```

```{r,fig.width=14}
makePlotTopN(gm.gathered,"Brazil",10)
```

```{r,fig.width=14}
makePlotTopN(gm.gathered,"South Africa",10)
```

```{r,fig.width=14}
makePlotTopN(gm.gathered,"Kenya",10)
```

```{r,fig.width=14}
makePlotTopN(gm.gathered,"Iraq",10)
```

Note: I wondered how @dpseidel had her Tidy Tuesday plots displayed on Github and discovered [this](https://rmarkdown.rstudio.com/github_document_format.html).
